\documentclass{article}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={210mm,297mm},
 left=20mm,
 right=20mm,
 top=-2mm,
 bottom=2mm,
 }
%\usepackage[margin=0.5in]{geometry}

\usepackage{amsmath,amssymb}
\usepackage{ifpdf}
%\usepackage{cite}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{pdfpages}
\usepackage{mdwtab}
\usepackage{eqparbox}
%\onecolumn
%\input{psfig}
\usepackage{color}
\usepackage{graphicx}
\setlength{\textheight}{23.5cm} \setlength{\topmargin}{-1.05cm}
\setlength{\textwidth}{6.5in} \setlength{\oddsidemargin}{-0.5cm}
\renewcommand{\baselinestretch}{1}
\pagenumbering{arabic}

\begin{document}

\textbf{
\begin{center}
{
\large{School of Engineering and Applied Science (SEAS),Ahmedabad University}\vspace{5mm}
}
\end{center}
%
\begin{center}
\large{Probability and Stochastic Processes (MAT277)\\ \vspace{4mm}
Homework Assignment-1\\\vspace{10mm}
Enrollment No: AU2140161 \hspace{3cm} Name: Mihir Kotecha } \vspace{2mm}
\end{center}}
\vspace{2mm}


\begin{enumerate}
\item \large 

\textbf{Alex told us that either face 1 or face 5 had turned up. Calculate the probability that
face 5 has turned up.}\\\\



\textbf{Let P(A) be the probability of getting a 1:} \\\\
$\therefore$ $P(A) = 0.24$\\\\\

\textbf{Let P(B) be the probability of getting a 5: } \\\\
$\therefore$ $P(B) = 0.04$\\\\

\textbf{It is given that either 5 or 1 has turned up, we need to find the probability of getting a 5.}\\\\

$\begin{aligned}
P(B|B \cup A) &= \cfrac{P(B \cap (B \cup A)}{P(B \cup A)} \\\\
&= \cfrac{P(B)}{P(A \cup B)}\\\\
&= \cfrac{P(B)}{P(A) + P(B)}\hspace{5mm} \textit{[Since A and B are mutually exclusive]}    \\\\
&= \cfrac{0.04}{0.04 + 0.24} \\\\
&= \cfrac{1}{7} = 0.143
\end{aligned}$

\newpage

\item \large 
\textbf{There are two events E1 and E2 and $P(E1 | E2) = 0.5$ and $P(E2 | E1) = 0.6$ and $P(E1 \cup E2)$ = 0.5}
\begin{enumerate}
    \item 
    \textbf{We need to find $P(E1 \cap E2)$: }\\\\
    $Given, \hspace{3mm}P(E1|E2) = \cfrac{P(E1 \cap E2)}{P(E2)} = 0.5$\\\\
    $\therefore$ \textit{from above equation: } $P(E2) = \cfrac{P(E1 \cap E2)} {0.5}$ \hspace{5mm} \textit{}{$---$ (1)}\\\\\\

    $Also, \hspace{3mm}P(E2|E1) = \cfrac{P(E1 \cap E2)}{P(E1)} = 0.6$\\\\
    $\therefore$ \textit{from above equation: } $P(E1) = \cfrac{P(E1 \cap E2)} {0.6}$ \hspace{5mm} \textit{}{$---$ (2)}\\\\\\

    \textbf{Now,}\\\\
    $P(E1 \cup E2)$ = 0.5\\\\
    $\therefore$ $P(E1) + P(E2) - P(E1 \cap E2) = 0.5$\\\\
    $\therefore$ $\cfrac{P(E1 \cap E2)}{0.5} + \cfrac{P(E1 \cap E2)}{0.6} - P(E1 \cap E2) = 0.5$\hspace{5mm} \textit{[From (1) and (2)]}\\\\
    $\therefore$ $P(E1 \cap E2)*(\cfrac{1}{0.5} + \cfrac{1}{0.6} - 1) = 0.5$\\\\\
    $\therefore$ $P(E1 \cap E2) = \cfrac{0.5}{2.67}$\\\\
    \boldmath $P(E1 \cap E2) = 0.187$


    \newpage \item
    \textbf{Now let us find if E1 and E2 are dependent or not: }\\\\

    \textit{On substituting the value of $P(E1 \cap E2)$ in (1): }\\\\
    $P(E2) = \cfrac{0.187}{0.5} = 0.374$\\\\\

    \textit{Similarly, on substituting the value in (2): }\\\\
    $P(E1) = \cfrac{0.187}{0.6} = 0.312$\\\\\

    \textbf{Now,}\\\\
    $P(E1)*P(E2) = 0.312 * 0.374 = 0.117$\\\\

    \textbf{We can see that the above value is not equal to $P(E1 \cap E2)$, thus the two events are Dependent. }
\end{enumerate}
            


\newpage

\item \large We are tossing an unfair dice repeatedly until we we get two heads in a row. The probability of getting a head $P(H) = p$. We need to find the winning probability. We win when we get two heads in a row. \\\\

The winning series can be as follows:

HH, THH, HTHH, THTHH, HTHTHH ... \\\\

$\begin{aligned}
    \therefore P(win) &= p^2 + (1-p)*p^2 + p*(1-p)*p^2 + p*(1-p)^2*p^2 + ....\\\\
    &= p^2*[1 + (1-p) + p*(1-p) + p*(1-p)^2 + ...]\\\\
    &= p^2*[(1 + p(1-p) + p^2(1-p)^2 + ...) + ((1-p) + p*(1-p)^2 + p^2*(1-p)^3 + ...)]\\\\
    &= p^2*[\cfrac{1}{1-p+p^2} + \cfrac{1-p}{1-p+p^2}]\\\\
    &= p^2*[\cfrac{2-p}{1-p+p^2}]
\end{aligned}$
\newpage

\item \large \textbf{A ten sided fair dice is rolled until a six is obtained. Let the number of throws be X.}
\begin{enumerate}
\item \textbf{We need to find the probability for X=5}\\\\
\textbf{For this to happen we need the six to occur exactly on the 5th throw}\\\\
$\therefore$ $P(X = 5) = \cfrac{9}{10}*\cfrac{9}{10}*\cfrac{9}{10}*\cfrac{9}{10}*\cfrac{1}{10}$\\\\

$\therefore$ $P(X = 5) = 0.065$\\\\\

\newpage

\item \large \textbf{Now we the conditional probability i.e. given $X>5$ what is $X \geq 6$: }\\\\
$\begin{aligned}
P(X \geq 6 | X>5) &= \cfrac{P(X\geq6 \cap X>5)}{P(X>5)}\\\\
&= \cfrac{P(X>5)}{P(X>5)}\\\\
&= 1
\end{aligned}$
	
\end{enumerate}

\newpage

\item \large \textbf{We are given the probabilities of arrival and leaving timings of employ. According to the question: }

\textbf{Let A be an event where the employ arrives late,}\\\\
$\therefore$ $P(A) = 0.15$\\\

\textbf{Let B be an event where the employ leaves early,}\\\\
$\therefore$ $P(B) = 0.3$\\\

\textbf{Also,}\\\\
$P(A \cap B) = 0.08$\\\\

\textbf{We need to find the probability of the employ arriving early given that he leaves late: }

$\begin{aligned}
P(A' |  B') &= \cfrac{P(A' \cap B')}{P(B')}\\\\
&= \cfrac{P(A \cup B)'}{P(B')}\\\\
&= \cfrac{(P(A) + P(B) - P(A \cup B))'}{P(B')}\\\\
&= \cfrac{1-(P(A) + P(B) - P(A \cup B))}{1 - (P(B))}\\\\
&= \cfrac{1 - (0.15 + 0.3 - 0.8)}{0.7}\\\\
&= 0.9
\end{aligned}$

\newpage

\item \large \textbf{We are given that A is an event in which the product breaks down in the third year and B is the event in which the product survives the first two years. Also, T is a random variable which represents the lifetime of a product.}\\\\

\textbf{Also,}\\
$P(T \geq t) = e ^ {\cfrac{-t}{5}} $ \textbf{, $t \geq 0$}\\\\
$P(T \geq 2) = 0673$\\\\

\textbf{We need to find the conditional probability A given B i.e. $P(A|B)$: }\\\\

$\begin{aligned}
    P(A|B) &= P(X=3 | X\geq2)\\\\
    &= \cfrac{P(X=3 \cap X\geq2)}{P(X \geq 2)}\\\\
    &= \cfrac{P(X=3)}{P(X \geq 2)}\\\\
    &= \cfrac{P(X \geq 2) - P(X \geq 3)}{P(X \geq 2)}\\\\
    &= \cfrac{e^\frac{-2}{5} - e^\frac{-3}{5}}{0.6703}\\\\
    &= \cfrac{0.6704 - 0.5488}{0.6703}\\\\
    &= \cfrac{0.1215}{0.6703}\\\\
    &= 0.1812
\end{aligned}$



\end{enumerate}

\end{document} 